### 2009-05-14 code/binding/development/specifications/bham_meeting/scenarios/yr3example/example.tex

\documentclass{article}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mdwlist}
\usepackage{graphicx}

\title{Example Specification Process for Year 3 PlayMate}
\date{\today}
\author{Nick and Henrik}

\begin{document}

\maketitle

\section{Script -- level 1}

\begin{enumerate}
\item \label{lvl1:redbox}
  human puts down red box
\item \label{lvl1:bluebox}
  human puts down blue box to left of right box
\item \label{lvl1:human}
  human says "put the blue thing to the right of the red thing"
\item \label{lvl1:success}
  robot does this successfully.
\end{enumerate}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{example1_msc.png}
\caption{
A sketch of the interaction between the involved systems.
}
\label{fig:lvl1}
\end{figure}

\section{Script -- level 2}

\begin{enumerate}

\item \label{lvl2:redbox}
  human puts down red box
\begin{enumerate}
\item \label{lvl2:redbox:notice}
    robot notices object,
\item \label{lvl2:redbox:extract}
    extracts colour and position information
\end{enumerate}

\item \label{lvl2:bluebox}
  human puts down blue box to left of right box
\begin{enumerate}
\item \label{lvl2:bluebox:notice}
    robot notices object,
\item \label{lvl2:bluebox:extract}
    extracts colour and position information,
\item \label{lvl2:bluebox:spatial}
    generates spatial relations between objects (blue object left of
    right object, blue object near red object, red object near blue
    object)
\end{enumerate}

\item \label{lvl2:human}
  human says "put the blue thing to the right of the red thing"
\begin{enumerate}
\item \label{lvl2:human:speech}
    robot recognises speech and creates a logical interpretation of it
\item \label{lvl2:human:thing}
    robot resolves the "thing" references in interpretation to the
    objects in the scene
\end{enumerate}

\item \label{lvl2:success}
  robot does this successfully.
\begin{enumerate}
\item \label{lvl2:success:goalgen}
    robot detects that the utterance is a command, and represents the
    the content of the command as a behaviour to perform.
\item \label{lvl2:success:plangen}
    robot generates a plan to pick up the blue object and move it to
    the right of the red object
\item \label{lvl2:success:manip}
    robot picks up the blue object and moves it to the right of the
    red object
\item \label{lvl2:success:verify}
    robot looks at the world to verify that the plan has been executed
    correctly
\end{enumerate}

\end{enumerate}

\section{Subarchitecture Level}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\linewidth]{example2_msc.png}
\end{center}
\caption{
A sketch of the interaction between the involved subsystems.
}
\label{fig:lvl2}
\end{figure}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:redbox:notice}}

\subsubsection{Visual Subarchitecture (VSA)}

\begin{itemize}

\item notification of scene change

\item representation of red box generated

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:redbox:extract}}

\subsubsection{VSA}

\begin{itemize}

\item colour value of red box

\item 3D position of red box

\end{itemize}

\subsubsection{Spatial Subarchitecture (SSA)}

\begin{itemize}

\item 2D position of red box on table

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:bluebox:notice}}

\subsubsection{VSA}

\begin{itemize}

\item notification of scene change

\item representation of blue box generated

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:bluebox:extract}}

\subsubsection{VSA}

\begin{itemize}

\item colour value of blue box

\item 3D position of blue box

\end{itemize}

\subsubsection{SSA}

\begin{itemize}

\item 2D position of blue box on table

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:bluebox:spatial}}

\subsubsection{SSA}

\begin{itemize}

\item left-of, blue object, red object

\item right-of, red object, blue object

\item near, blue object, red object

\item near, red object, blue object

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:human:speech}}

\subsubsection{Communication Subsubsection (CSA)}

\begin{itemize}

\item recognised speech string

\item logical form representing speech string

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:human:thing}}

\subsubsection{CSA}

\begin{itemize}

\item linguistic representation of ``red thing''

\item linguistic representation of ``blue thing''

\end{itemize}

\subsubsection{Binding Subarchitecture (BSA)}

\begin{itemize}

\item union binding together linguistic and visual representations of
  the red box

\item union binding together linguistic and visual representations of
  the blue box

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:success:goalgen}}

\subsubsection{BSA}

\begin{itemize}

\item Representation of command structure including arguments.

\end{itemize}

\subsubsection{Mediation Subarchitecture (MSA)}

\begin{itemize}

\item Description of the desired state of the world: right-of, blue
  object, red object

\item Goal to achieve this state

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:success:plangen}}

\subsubsection{MSA}

\begin{itemize}

\item Motivation to achieve goal state

\end{itemize}

\subsubsection{Planning Subarchitecture (PSA)}

\begin{itemize}

\item Plan to pick up blue thing, place it to the right of red thing

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:success:manip}}

\subsubsection{Manipulation Subarchitecture (ASA)}

\begin{itemize}

\item Actions to pick up blue box, place down at position to right of
  red box.

\end{itemize}

%% ------------ next script line ------------ %%

\subsection{\ref{lvl2:success:verify}}

\subsubsection{VSA}

\begin{itemize}

\item New position information for blue box

\end{itemize}

\subsubsection{SSA}

\begin{itemize}

\item new 2D position of blue box

\item right-of, blue object, red object

\item left-of, red object, blue object

\item near, blue object, red object

\item near, red object, blue object

\end{itemize}

\subsubsection{PSA}

\begin{itemize}

\item Representation of current state and desired state, and
  comparison

\end{itemize}

\section{Dependency Level}

What else is required to generate the above information and behaviour.

\section{Component Level}

First pass as describing the behaviours in terms of the components
that are involved.

\section{Responsibility Level}

Who's doing what!

\end{document}

% LocalWords:  CSA MSA PSA