### 2010-11-16 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
More paper drafting.
 - Charles

% File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}

\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\include{macros}

\nocopyright

\pdfinfo{
/Title (Switching in Continual Planning for Practical Robot Control)
/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
/Author (NA)}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
%\title{Switching in Continual Planning for Practical Robot Control}
\title{A Continual Planner that Switches Between Serial and
  Decision-Theoretic Planning} 
\author{NA}
\setcounter{secnumdepth}{0}

\begin{document} 
\maketitle

\begin{abstract}

The problem of practical robot control poses many important challenges
to automated planning.

Planning and plan execution must be robust, seamlessly accommodating
exogenous events and the underlying unpredictability of the
environment.

Plans with a reasonable chance of exhibiting good behaviours must be
synthesised in a timely manner.

interleaved planning and execution.

, reason about degrees of belief and uncertainty about the world,

\end{abstract}

\section{Introduction}

We describe, \pcogx, the planning component in a robotic system that
continuously deliberates in a stochastic dynamic environment in order
to achieve objectives set by the user, and acquire a good
understanding of its environment.
%%

Have a good understanding of the contents and function of rooms, as
well as the linguistic referrents to rooms, widgets, and their visual
qualities.

Having confidence in its beliefs about the linguistic terms for
spaces, and the function of those spaces.

sometimes difficult to predict, with exogenous events, such as the
changing of the objective,

the speed and scalability of software for serial planning in
deterministic

We developed a first-order declarative language, called DTPDDL, for
describing domains of planning problems that correspond to POMDPs.  To
accompany that language we have developed an information-state
\laostar\ procedure for solving problems expressed in DTPDDL. We have
an automated determinisation procedure that generates MAPL
descriptions given DTPDDL input. To accompany that, we have extended
MAPSIM to simulate DTPDDL instances, and modified \fastdownward\ so
that it can find useful serial executions given DTPDDL and MAPL models
of the problem at hand.

 called DTPDDL, along
the lines of PDDL for the partially observable case, an \laostar\
solution procedure, a determinisation of the DTPDDL problem in MAPL,
and modify the \fastdownward\ system to find high quality serial plans
.

We model the environment as a partially observable Markov decision
process. Although planning in that model is undecidable in
general~\cite{}, an optimal finite-horizon plan corresponds to a
contingent plan, that is a function mapping action-observation
histories to actions.

our continual planner is reactive, replanning whenever the underlying
domain and problem models change. For example, replanning occurs if
the motivational component alters the objectives, and if an assumed
outcome of a sensing action is not realised.

brittle sensing model,

the evaluation of a fluent at a state is either known. Moreover, there
is a sensing process that run-time variables  after-which

%% For $\prop \in \state$ we say proposition $\prop$ characterises state
%% $\state$. There is always a unique starting state $\state_0$. The goal
%% $\goal$ is a set of propositions, and we say that state $\state$ is a
%% goal state iff $\goal \subseteq \state$.

%%  To keep this exposition
%% simple, for any two distinct actions $\stochAction_i \neq
%% \stochAction_j$, if outcome $\detAct$ is a possibility for
%% $\stochAction_i$ then it cannot also be a possibility for
%% $\stochAction_j$ -- i.e., if $\detAct \in
%% \detActions(\stochAction_i)$ then $\detAct \not\in
%% \detActions(\stochAction_j)$.

%% The solution to a probabilistic planning problem is a contingency
%% plan. This consists of an assignment of actions to states at each
%% discrete timestep up to the planning horizon $n$. The optimal
%% contingency plan is one which prescribes actions to states that
%% maximise the probability that the goal is achieved within $n$ steps
%% from the starting state $\state_0$. For the purposes of this paper
%% we say a plan fails, i.e. achieves the goal with probability $0$, in
%% situations where it does not prescribe an action. Computing the
%% optimal plan for a problem is computationally intractable, and an
%% important direction for research in the field is to develop
%% heuristic mechanisms for generating small linear plans
%% quickly~\cite{littman:etal:98}

An outline of the paper is as follows. We describe POMDPs with a
propositionally factored representation of states and observations,
and describe how to evaluate plans when they correspond to
finite-state controllers.

\section{Propositional Decision-Theoretic Planning}

%% {\em runtime variables}
%% {\em decision variables}
%% {\em uninitialised variables}
%% {\em omitted variables}

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. The underlying world
dynamics are modelled in terms of a finite set of stochastic actions
$\stochActions$, deterministic outcomes $\detActions$, and
state-characterising propositions $\propositions$.  A problem state
$\state$ is a set of propositions $\state \subseteq \props$, and we
denote \states\ the set of reachable states. The perceptual model is
given in terms of a finite set of stochastic senses $\stochSenses$,
deterministic sensing outcomes $\detSenses$, and perceptual
propositions $\percepts$, called percepts. Here, an observation
$\observ$ is a set of percepts $\observ \subseteq \percepts$, and we
denote \observations\ the set of reachable observations.

Every stochastic action $\stochAction \in \stochActions$ has a
precondition $\poss(\stochAction)$, which is a set of propositions. An
action is applicable at a state $\state$ when $\poss(\stochAction)
\subseteq \state$. We denote $\stochActions(\state)$ the set of
actions applicable at state $\state$.  When $\stochAction \in
\stochActions(\state)$ is applied, nature chooses one elements amongst
a small set of deterministic outcomes $\detActions(\stochAction)
\equiv \{\detAction_1, \ldots, \detAction_k \}$. We denote
$\mu_{\stochAction}(\detAction_i)$ the probability that nature takes
outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. That outcome so chosen has an
effect on the underlying state given in terms of two lists of
propositions. The add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
chosen by nature when \stochAction\ is applied at state $\state$, then
the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
added to $\state$, and those from $\delete(\detAction)$ are removed
from $\state$.

We are concerned with problems that feature partial
observability. Here the underlying state of the world cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in \observations$ that informs what
should be believed about the world. In the POMDP model, if $\action$
is applied effecting a transition to a successor state $\state'$, then
an observation occurs according to the active senses
$\stochSenses(\stochAction, \state') \subseteq \stochSenses$. A
precept is active, written $\stochSense \in
\stochSenses(\stochAction, \state')$, if the senses
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition
$\poss_\states(\stochSense) \subseteq \propositions$ is satisfied by
the state $\state'$; In the usual sense that
$\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

A POMDP has a starting configuration that corresponds to a subjective
Bayesian belief state. Written $\bstate_0$, this is a probability
distribution over states. We use the notation $\bstate_0(\state)$ to
denote the probability that the starting state is equal to
$\state$. We suppose this is given in a factored tree format.

%% Let $\detSense_i \subseteq \percepts$ for each $i$, so that if
%% we apply $\stochAction$ and transition to $\state'$, the observation
%% received will be a union of choices from the active percepts, which
%% will occur with the probability according to $\psi$ that nature can
%% make those choices.

\subsection{Costs, Rewards, and Plan Evaluation}

Whereas until now we have considered a POMDP model factored in terms
of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of states, \actions\ is the
finite set of actions, and \observations\ is the finite set of
observations.  Where $\state, \state' \in \states$, $a \in \actions$,
from $\mu$ we have a state transition function $\transProb(\state,
\action, \state')$ giving the probability of a transition from state
$s$ to $s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function.  Consequently there is a positive constant $c$
so that for all $\state \in \states$ and $\action \in \actions$,
$|\reward(\state, \action)| < c$.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

A general formalism for representing solutions to POMDPs is the
finite-state controller (FSC). This is a three-tuple $\langle \nodes,
\selection, \transition \rangle$ where: $\node \in \nodes$ is a set of
nodes, $\selection_\node(\action) = P(\action | \node)$, and
$\transition_\node(\action, \observation, \node') = P(\node'|\node,
\action, \observation)$. The value of state $\state$ at node $\node$
given a problem model, written $V_\node(\state)$, is:

\begin{equation}\label{eq:evaluation}
V_\node(\state) = \sum_{\action \in \actions}
\selection_\node(\action) \reward(\state, \action) + \beta \sum_{\action, \observation,
\state', \node'} \transition_\node(\action, \observation, \node')
\transProb(\state, \action, \state') \obsDist_\observ(\state',
\action) V_{\node'}(\state')
\end{equation}

Using the notation \bstate\ for a belief-state, and writing
$\bstate(\state)$ for the probability the robot believes it is in
state \state, we have that the value of that belief $\bstate$ according to a
controller is:

\begin{equation} \label{eq:valueBelief}
V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
\end{equation}

\section{Serial Planning in Determinisations}

\section{Decision-Theoretic PDDL}

the effects of a sensing schema are perceptual, whereas the effects of
an operator schema are on the underlying state space.

knowledge objectives and  sensory effects.

probabilistic models of the sensing consequences of acting to
quantitatively capture the unreliability of sensing, and thus reason
precisely about the utility of a plan.

planning in practical sized POMDPs is intractable. One approach is to
described the task hierarchically, thereby constrain and simplify the
policy search task. The approach ~\cite{}.

It can arise that in some propositional states an action is
applicable, and in others it is not. Moreover, a belief-state can
assign non-zero probability to states where that applicability holds,
and states where it does not. We forbid execution of actions that are
not applicable in any state of the current belief, and otherwise
suppose that when a non-applicable action is executed, that it has
no effect.

\section{Evaluation}

\section{Discussion}

There have been a number of developments recently towards planning
under uncertainty using systems that were intended for serial planning
in deterministic problems.  Notably,
\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
probabilistic track of the 2004 International Planning Competition,
uses the fast satisficing procedure
\system{FF}~\cite{hoffmann:nebel:2001} to compute a serial plan and
corresponding execution trace. The resulting plan is then executed
until the observed trace deviates from the planned trace.

 ---i.e., the values of {\em
runtime variables} in the continual paradigm--- are treated , and
therefore plan for a specific eventuality, replanning from scratch if
something

More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures. The outcome is typically a more
robust serial plan~\cite{yoon:etal:2008}, or contingent
plan~\cite{majercik:2006}.

%%
Recently the system
\system{Conformant-FF}~\cite{hoffmann:brafman:2006}, demonstrated how
conformant planning ---i.e., serial planning in unobservable worlds---
can be modelled as a deterministic problem, and therefore solved using
serial systems. In this conformant setting, advances have been towards
compact representations and lazy evaluations of belief-states.

\section{Concluding Remarks}

The switching continual planning system we have described serves as
the underlying planner for CogX.

\bibliography{papers}
\bibliographystyle{aaai}
\end{document}

\subsection{PPDDL}

We briefly discuss the description language PPDDL, a fairly
straight-forward extension of the Planning Domain Definition Language
for describing fully observable Markov decision processes.

\small{
\begin{tabtt}
(\= :action look-at-object \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - label ?p - place) \\
  \> :precondition (and (= (is-in ?r) ?p) \\
  \> \> (= (label ?o) ?l) ) \\
  \> :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
}

\small{
\begin{tabtt}
(\= :observe visual-object \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - label ?p - place) \\
  \> :execution (look-at-object ?r ?v ?l ?p) \\
  \> :effect (when (= (is-in ?o) ?p) (probabilistic 0.8)) \\
\end{tabtt}
}

\small{
\begin{tabtt}
(\= :sensor look-for-object \\
  \> :agent (?a - robot) \\
  \> :parameters (?l - label ?p - place) \\
  \> :variables (?o - visual-object) \\
  \> :precondition (\=and (= (is-in ?r) ?p) \\
  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?p) ) \\
  \> :effect () \\
  \> :sense (= (is-in ?o) ?p) \\
\end{tabtt}
}

\subsection{MAPL}