### 2010-11-23 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/discussion.tex
New section

A theoretical criticism of switching continual planning, concerns
interleaved sequential and decision-theoretic sessions failing to make
any progress towards the objective.