### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
got new draft of switching, and Bayesian stuff for the evaluation section

% coming soon

Whether proceeding in a sequential or contingent session, our
continual planner must maintain a factored representation of the
current underlying belief-state, by performing belief revision
according to Eq~\ref{eq:revision}. That distribution corresponds to
the source of determinisations for sequential planning, the basis for
switching, and as the source for state abstraction in contingent
sessions. In our implementation we have
used~\citeauthor{king:2009}~(\citeyear{king:2009}) for this purpose.