### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex
Few updates

% coming soon

In our implementation uses the
~\citeauthor{king:2009}~(\citeyear{king:2009}) \system{Dlib-ml} for
successive estimation of the underlying belief-state.