### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/evaluation.tex

% coming soon

We have implemented our switching continual approach in the MAPSIM
environment~\cite{brenner:nebel:jaamas09}. Our implementation is able
to use several underlying planning systems. We have also extended
MAPSIM system so that it can parse DTPDDL, and perform successive
estimation of the underlying belief-state using
\system{dlib-ml}~\cite{king:2009} for inference.  In this evaluation
we use our own version of Fast Downward~\cite{fast-downward} for
sequential sessions. We have extended that system to support actions
with success probabilities. In our evaluation Fast Downward is run
with the cyclic causal graph heuristic using an A* search or weighted
A* (with weight 5). We use A* for easy problems and WA* for more
difficult problems where A* is ineffective. Contingent sessions use
our own forward search procedure.

\Omit{ We then perform multiple tests with different limits for the
  belief-space size in contingent sessions.  Higher limits should
  cause longer planning times but be beneficial to plan quality as
  more contingencies can be taken in to account by the POMDP planner.
}

In order to evaluate our system, we have also implemented a baseline
approach in MAPSIM. Rather than invoking a contingent session when a
switching action $\switchAction$ is scheduled for execution, the
baseline simply executes one action that can trigger a sensing outcome
determined by the precondition of $\switchAction$, and replans in the
resulting belief-state.

%% executes switching actions and replans in the
%% resulting belief state.

%% but instead of
%% creating an observation problem for the decision theoretic planner it
%% will just execute one sensing action -- assuming that this action will
%% confirm its assumption.

To test our approach we use a robot exploration domain based on a that
from our physical robotic system. Here, a robot is exploring an office
or living environment, and trying to report the locations of objects
to their owner.

\begin{figure}[h]
  \centering
  \input{figures/dora}
  \caption{A exploration domain with three rooms and 13 places. The
    robot is in in the centre of room A}
\label{fig:dora2}
\end{figure}
An instance of this domain is shown in figure \ref{fig:dora2}. Basic
types of element are{\tt rooms}, {\tt places} and {\tt
  objects}. Places are topological map nodes that occur in rooms.
Objects, including the robot, are always located at a place. The robot
can move around the rooms via connections between places given by the
{\tt connected} predicate. Each room has a, possibly unknown, {\em
  category} (e.g. kitchen, office, living room) and depending on this
category, objects are placed inside the rooms.

The robot can find out if an object is at a certain place by executing
the {\tt look-for-object} action. Moreover, the contents of a room
indicates its category, and certain objects are more likely to be
located in rooms of certain categories. For example, a box of
cornflakes is more likely to be located in the kitchen than the
office. The {\tt look-for-object} might result in a positive
perception if the object is there. Some objects are harder to detect
than others, and an instantaneous positive detection of the object is
not proof positive that it is there. Additionally, if the robot is in
the presence of a human, it can ask what type of room they are
currently in, however conducting a dialogue is more costly than
running a vision algorithm (cost of 8 vs costs of 3).

We perform experiments on tasks of several sizes. The robot's
objective is to find one or more objects and report back their
position to a person. In order to determine the impact of sensor
reliability in a comparison with the baseline, we also run tests on
identical problems but changed the sensor model for the target
objects: the probability for perceiving the object if it was there
went from 0.9 in the easiest case to 0.65 and 0.4 in the average and
hard cases respectively. In addition, we test how our system performs
in tasks requiring indirect sensing. Here, we gave the planner rewards
for visiting a certain type of room (e.g. kitchen or office). We did
not place any humans in those scenarios, so the only way of
determining the room category is by looking for objects typical for
that room type. As this indirect reasoning is a tasks that cannot be
performed by the continual planner alone, we did not run the baseline
system on those problems.

In our evaluation we run 50 simulations in each configuration. The
planning and simulation models are in exact correspondence. Not all
problems have a solution, so we also examine whether or not detection
of the non-existence of a solution is important in stochastic domains.
%% 
We perform multiple tests with different limits for the
belief-space size in contingent sessions.  Higher limits result in
longer contingent planning times and higher plan quality.

Using a satisficing, optimistic serial planner in sequential sessions
makes optimising for the expected reward difficult. Only reasoning
about single traces, it is optimistic wrt the remaining costs to a
goal and pessimistic regarding the probability of reaching tat. By
setting the rewards to non-extremal values has little effect on that
system. Therefore, we do not consider expected utility in our
experiments. Instead, we report average costs of the plans and the
success rate (as a ratio between solved and solvable problems).

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-time}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-time}\hfill
  \vspace{2mm}
  \includegraphics{dora3-time}\hfill
  \vspace{2mm}
  \includegraphics{dora4-time}\hfill
  \vspace{2mm}
  \includegraphics{dora56-time}\hfill
  \vspace{2mm}
  \includegraphics{dora-cat-time}\hfill
  \caption{Average runtime}
  \label{fig:results-time}
\end{figure}

\begin{figure}[h!]
  % \centering
  % \includegraphics{dora1-quality}\hfill
  % \vspace{2mm}
  \includegraphics{dora2-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora3-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora4-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora56-quality}\hfill
  \vspace{2mm}
  \includegraphics{dora-cat-quality}\hfill
  \caption{Average plan costs and number of successful runs}
  \label{fig:results-quality}
\end{figure}

%% Figure \ref{fig:results-quality} shows the average costs of the
%% executed plans as well as the percentage of solvable tasks that were
%% actually solved by the planner.

Our experimental results are summarised in
Figure~\ref{fig:results-time} and Figure~\ref{fig:results-quality}. We
find that the time spent on contingent planning increases steeply as
the abstraction $\bstate_0$ becomes more refined. For coarse initial
configurations, the cost of contingent planning is compensated by the
decrease of time spent in Fast Downward. Addressing success ratios and
plan costs, for objects that can be easily detected there is little
gain in using a decision-theoretic planner, as the greedy sensing
approach by the baseline is sufficient. With decreasing sensor
reliability contingent planning pays off.%% : while the resulting plans are still
%% longer on average, the impact on the number of solved tasks was much
%% smaller than for the baseline system.  Less aggressive abstraction
of the initial configuration results in longer runtimes. In our
scenario it is worthwhile being assumptively aggressive.  We suppose
the relatively high success rate irrespective of the coarseness of the
initial configuration indicates the effectiveness of using conditional
entropy to guide abstraction refinement.

We find that abstraction has little impact on plan
costs and success rate in our scenario. Increasing the size of the
initial abstract belief beyond 50 states rarely pays off, because
while additional information may facilitate better plans, the indirect
sensing tests report the same result. Here, we only really see an
impact in terms of the runtime of contingent planning.

% We believe that a part of the improvement is due to the segmentation of
% the plan into several subtask, essentially performing hierarchical
% planning. Especially when the continual planner performs badly this is
% a huge gain.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "moritz_2011"
%%% End: