### 2011-01-28 subarchitectures/planner.sa/trunk/src/latex/AAAI11/aaai11.tex
AAAI version of the switching planner paper

% File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}

\usepackage{subfig}

\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{positioning}

\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{graphicx}
\graphicspath{{plots/}}
\include{macros}%

\nocopyright

\pdfinfo{
/Title (Switching in Continual Planning for Practical Robot Control)
/Subject (Proceedings of the 21st International Conference on Automated Planning and Scheduling)
/Author (NA)}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
%% \title{Interleaved Serial and Decision-Theoretic Planning for a Mobile Robot}

%% \title{Interleaved Serial and Decision-Theoretic Planning for
%% Practical Robot Control}

\title{Interleaved Serial and Decision-Theoretic Planning for High-Level Robot Control}

%% \title{A Switching Planner for Combined Task and Observation Planning}
\author{anonymous}
\setcounter{secnumdepth}{0}

\begin{document} 
\maketitle

\begin{abstract}

%% Those processes must

From an automated planning perspective the problem of practical mobile
robot control in realistic environments poses many important and
contrary challenges.  On the one hand, the planning process must be
lightweight, robust, and timely. Over the lifetime of the robot it
must always respond quickly with new plans that accommodate exogenous
events, changing objectives, and the underlying unpredictability of
the environment.  On the other hand, in order to exhibit useful
behaviours the planning process must perform computationally expensive
reasoning about contingencies and possible revisions of subjective
beliefs according to quantitatively modelled uncertainty in acting and
sensing.
%%
Towards addressing these challenges, we develop a \emph{continual
planning} approach that {\em switches} between using a fast
satisficing ``classical'' planner, to decide on the overall strategy,
and decision-theoretic planning to solve small subproblems where
deeper consideration of the sensing model is both practical, and can
significantly impact overall performance. We demonstrate our approach
in large problems from a realistic robot exploration domain.

%% partial observability

%% In cooperation with execution monitoring, it must perform repair
%% and/or replanning in order to seamlessly accommodate exogenous events,
%% changing objectives, and the underlying unpredictability of the
%% environment.

%% Realistic robot planning problems in uncertain environments often
%% require achieving tasks while also finding out about the
%% world. Because the world state cannot be observed directly, these
%% problems are naturally represented as partially observable Markov
%% decision problems (POMDPs). However, these are typically intractable
%% for realistic problems.
%%

\end{abstract}

%% \section{Moritz's Initial Outline}

%% \subsection{Switching Planner Outline}

%% Outline the high level approach here:
%% \begin{itemize}
%% \item Problem input (factored initial state, DTPDDL observations)
%% \item Determinisation
%% \item Problem generation for DT
%% \item Belief revision
%% \end{itemize}

%% \subsection{Determinisation}

%% \begin{itemize}
%% \item Generation of deterministic sensing models
%% \item Determinisation of the initial state
%% \item Compiling reward model into (soft-)goal model
%% \item Cost function in FD (maybe this could mentioned in the outline?)
%% \end{itemize}

%% \subsection{Subtask Generation}
%% \begin{itemize}
%% \item Goal generation from CP trace
%% \item Disconfirm actions
%% \item State pruning according to CP trace
%% \end{itemize}

%% \subsection{Belief revision}
%% Not sure if this needs its own section
%% \begin{itemize}
%% \item Convert inital state into bayesian network
%% \item Belief updates with observation nodes
%% \item Create new network for next planning runs
%% \end{itemize}

\section{Introduction}
\input{intro.tex}

\section{Propositional Decision-Theoretic Planning}
\input{prop-dtp.tex}

\section{Decision-Theoretic PDDL}
\input{dtpddl.tex}

\section{Switching Continual Planner}
\input{switching.tex}

\section{Evaluation}
\input{evaluation.tex}

%% \section{Discussion}
%% \input{discussion.tex}

\section{Related Work}
\input{relatex.tex}

\section{Concluding Remarks}
\input{conc.tex}

\bibliographystyle{aaai}
\bibliography{papers}

\end{document}

\subsection{PPDDL}

We briefly discuss the description language PPDDL, a fairly
straight-forward extension of the Planning Domain Definition Language
for describing fully observable Markov decision processes.

\small{
\begin{tabtt}
(\= :action look-at-object \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - label ?l - location) \\
  \> :precondition (and (= (is-in ?r) ?l) \\
  \> \> (= (label ?o) ?l) ) \\
  \> :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
}

\small{
\begin{tabtt}
(\= :observe visual-object \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - label ?l - location) \\
  \> :execution (look-at-object ?r ?v ?l ?l) \\
  \> :effect (when (= (is-in ?o) ?l) (probabilistic 0.8)) \\
\end{tabtt}
}

\small{
\begin{tabtt}
(\= :sensor look-for-object \\
  \> :agent (?a - robot) \\
  \> :parameters (?l - label ?l - location) \\
  \> :variables (?o - visual-object) \\
  \> :precondition (\=and (= (is-in ?r) ?l) \\
  \> \> (= (label ?o) ?l) (assume (is-in ?o) ?l) ) \\
  \> :effect () \\
  \> :sense (= (is-in ?o) ?l) \\
\end{tabtt}
}

\subsection{MAPL}

%% \scriptsize
%% \begin{nopagebreak}\begin{tabtt}
%% <observation-def> \=::= \=(:sense <observation symbol> \\
%%                         \> \> :parameters (<typed list (variable)>)  \\
%%                         \> \> <o-def body>) \\
%%   <o-symbol> \> ::= <name> \\
%%   <o-def body> \> ::= [:precondition <GD>] \\
%%   \> \> [:execution <atomic action(term)> ] \\
%%     \> \> [:effect <o-effect>] \\
%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
%%   <o-effect> \> ::= <c-o-effect> \\
%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
%%   <c-o-effect> \> ::= <p-o-effect> \\
%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
%%   <o-f-exp> \> ::= <number>\\
%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
%%   <o-f-exp> \> ::= <o-f-head>\\
%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
%%   <o-f-head> \> ::= <o-function-symbol>\\
%% \end{tabtt}\end{nopagebreak}
%% \noteme{<structure-def> ::= <attach-def>}
%% \normalsize

%% \footnote{We have simplified the schemata for illustrative
%% purposes. In practice, the probabilities of a particular observation
%% here should be parametrised by the type of visual object the robot is
%% searching for, and category of the location.}

We suppose $\bstate_0$ is given in a factored tree format of the
form:

%% \small{
%% \begin{tabtt}
%% (:init  (and \=f_{11} f_{12} ..
%%  \> (probabilistic p_{}) 
%%  \> (probabilistic ..) ..) \\
%% \end{tabtt}
%% }

%% \[(:\pp{init} (and f_{11} f_{12} .. (\pp{probabilistic} p_) ))\]

planning in practical sized POMDPs is intractable. One approach is to
described the task hierarchically, thereby constrain and simplify the
policy search task. The approach ~\cite{}.

%% More generally, a useful
%% formalism for representing solutions to POMDPs is the finite-state
%% controller (FSC). This is a three-tuple $\langle \nodes, \selection,
%% \transition \rangle$ where: $\node \in \nodes$ is a set of nodes,
%% $\selection_\node(\action) = P(\action | \node)$, and
%% $\transition_\node(\action, \observation, \node') = P(\node'|\node,
%% \action, \observation)$. Where the value of acting is the discounted
%% accumulated reward over an infinite horizon, then each controller can
%% be assigned a value, written $V_\node(\state)$, expressed in the usual
%% way according to the Bellman equation:

%% \begin{equation}\label{eq:evaluation}
%% \begin{array}{lcl}
%% V_\node(\state) & = & \sum_{\action \in \actions}
%% \selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\

%% && \hspace{-10ex} \beta \sum_{\action, \observation,
%% \state', \node'} \transition_\node(\action, \observation, \node')
%% \transProb(\state, \action, \state') \obsDist_\observ(\state',
%% \action) V_{\node'}(\state')
%% \end{array}
%% \end{equation}

%% \noindent Here, the value of {\em prior} $\bstate$ given an FSC is
%% then:

%% \begin{equation} \label{eq:valueBelief}
%% V_{\pp{FSC}}(\bstate) = \max_{\node \in \nodes} \sum_{\state \in \states} \bstate(\state) V_\node(\state)
%% \end{equation}

%% \noindent There is a corresponding deterministic FSCs for both
%% sequential (i.e., conformant) and contingent
%% plans. Eq~\ref{eq:evaluation} gives their value in the finite-horizon
%% case, if we take $\beta = 1$, and alter the process dynamics so there
%% is a compulsory transition to a zero utility sink state when that
%% horizon is reached.

 ---i.e., the values of {\em
runtime variables} in the continual paradigm--- are treated , and
therefore plan for a specific eventuality, replanning from scratch if
something

Sequential planning in our approach proceeds in two phases. In the
first phase, the goal is to identify a plan prefix

In sequential planning, the plan prefix corresponds to choosing a
valid abstract starting state.

moreover, for the purpose of decision-theoretic abstractions,

A relaxed form of {\em visitation} we call {\em abstract
  visitation}. In the abstract case a conjunctive term is visited
  provided all its atomic subterms are visited, and zero or more of
  its probabilistic terms are visited.

The planner has a compulsory starting action, that visits the root
term, and then has to execute visitation actions until that plan
prefix corresponds to a valid {\em abstract visitation} of the {\em
:init} decleration. Beyond such a prefix, the process actions and senses are
available to the planner

Goal expressions in the problem description ---for example
$(\pp{kval}~(\pp{location}~\pp{box}))$ expresses ``the robot knows the
location of the {\em box}--- are treated as control-knowledge by the
planner. Here, a plan is not considered valid unless in the finial
stet the goal condition is satisfied.