### 2011-02-03 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/relatex.tex
New bibliography, and the related work section is begenning

\cite{wyattetal2010tamd}

For the purpose of symbolic
planning,~\cite{wyattetal2010tamd,talamadupula:2010,Kraft2008} treat
sensing deterministically and beliefs qualitatively.
%%
 Addressing specifically this second
challenge,~\citeauthor{talamadupula:2010} identify
\emph{continual planning} in the presence of detailed probabilistic
 models as an important direction for future research.

Focusing now specifically on planning, our work is motivated by
domains that contain a mixture of task planning and observation
planning. There have been a number of recent papers representing
observation planning problems as POMDPs and using various techniques
to manage the large state
space. \citeauthor{hippo-jnl}~(\citeyear{hippo-jnl}) take this
approach in a vision algorithm selection problem. In their case there
is a natural hierarchical decomposition of the problem which allows
them to solve large problems by breaking them into a set of small
POMDPs. \citeauthor{doshi08:pref_elic}~(\citeyear{doshi08:pref_elic})
represent a preference elicitation problem as a POMDP and take
advantage of symmetry in the belief space ---essentially the idea that
it does not matter what the value of the variable you are trying to
observe is--- to exponentially shrink the state space. Although we
have been actively exploring the \citeauthor{doshi08:pref_elic}
approach, those exploitable structures are not present in our domain
due to the task planning requirement.

\cite{shani:etal:08}

Generally There has been much recent work on scaling POMDP solution procedures
to medium-sized instances. In the case of general domain-independent
factored systems, the state-of-the-art scales to relatively small
problems with $2^{22}$
states~\cite{shani:etal:2008}.\footnote{Considering only room
categories and distribution of objects, problems we consider in this
paper have $\sim 10^{27}$ states. The details of view points, from
local active visual search, and those of robot location, further
increase that figure. Therefore, not only because they are offline,
but also because they have limited scalability, these approaches are
infeasible in our setting.} At their limit, these procedures take over an
hour to converge.  For classes of POMDP that feature exploitable
structures (e.g., no actions with negative effects), problems with as
many as $10^{30}$ states can be targeted by offline
procedures~\cite{brunskill:russell:2010}. Moving someway towards
addressing all the challenges we have outlined, recent online POMDP
solution procedures have been developed which can exploit highly
approximate value functions -- typically computed using a point-based
procedure -- and heuristic in forward
search~\cite{ross:etal:2008}. Their applicability in our setting is
limited due to the large amount of
\emph{problem-specific} offline processing required to get useful
search guidance. A {\em very} recent and promising online approach for
large POMDPs employs Monte-Carlo sampling to break the curse of
dimensionality in situations where goal reachability is
easy~\cite{silver:veness:2010}. Although we suppose it an interesting
item for future work to pursue that direction, it should be noted that
ease of goal reachability is not guaranteed in the problems we face,
and is certainly not a property to be assumed in domain independent
planning.

There have been a number of recent papers on
planning under uncertainty using systems that were intended for
sequential planning in deterministic problems,  for example,
\system{FFR$_a$}~\cite{yoon:etal:2007}, the winning entry from the
probabilistic track of the 2004 International Planning Competition.
In the continual paradigm, \system{FFR$_a$} uses the fast satisficing
procedure \system{FF}~\cite{hoffmann:nebel:2001} to compute sequential
plans and corresponding execution traces.
%%
More computationally expensive approaches in this vein combine
sampling strategies on valuations over {\em runtime variables} with
deterministic planning procedures. The outcome is typically a more
robust sequential plan~\cite{yoon:etal:2008}, or contingent
plan~\cite{majercik:2006}.

%%No! They simply haven't been evaluated in PO settings. They may, or
%%may not struggle. They have sampling of traces, and that would
%%include observations, and therefore evolutions of beliefs. SSAT was
%%proposed by Littman for POMDPs. So the majercik stuff is perfectly
%%suited to POMDPs.

% Normally if we are going to compare with related work, we do
% actually *compare*. Why didn't you try those approaches? I think
% it's safe to say that FFR will struggle. Why would it even include
% in its plan an observational action that doesn't change the world?

%%
%% However, as we said in the introduction,
%% all these approaches struggle in partially observable domains as they
%% rely on being able to determine the state at all times.

Also leveraging deterministic planners in problems that feature
uncertainty, \system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
$T_0$~\cite{palacios:geffner:2009} demonstrate how conformant planning
---i.e., sequential planning in unobservable worlds--- can be modelled
as a deterministic problem, and therefore solved using sequential
systems. In this conformant setting, advances have been towards
compact representations of beliefs amenable to existing best-first
search planning procedures, and lazy evaluations of beliefs. Most
recently this research thread has been extended to contingent planning
in fully observable non-deterministic
environments~\cite{albore:etal:2009}.
%%
The continual planning system that motivated our
project~\cite{brenner:nebel:jaamas09} also has this characteristic,
and has been applied in completely observable domains, particularly
those featuring multiple communicating agents.

%% The use of knowledge
%% operators in domains allows plans that act to gain knowledge, but the
%% approach assumes that such actions are deterministic and reliable, an
%% assumption that we relax.