### 2011-04-21 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex

\newcommand{\entropy}{\ensuremath{\mathrm{H}}}

We now describe our {\em switching} planning system that operates
according to the continual planning paradigm. The system {\em
switches} in the sense that planning and plan execution proceed in
interleaved sessions in which the {\em base planner} is either {\em
sequential} or {\em decision-theoretic}.  The first session is
sequential, and begins when a DTPDDL description of the current
problem and domain are posted to the system.  During a sequential
session a serial plan is computed that corresponds to one
execution-{\em trace} in the underlying decision-process. That trace
is a reward-giving sequence of process actions and {\em assumptive}
actions. Each {\em assumptive} action corresponds to an assertion
about the truth of a fact that is unknown at plan time -- e.g. that a
box of cornflakes is located on the corner bench in the kitchen. The
trace specifies a plan and characterises a {\em deterministic
approximation} (see~\cite{yoon:etal:2008}) of the underlying process
in which that plan is valuable. Traces are computed by a
cost-optimising classical planner which trades off action costs, goal
rewards, and determinacy. Execution of a trace proceeds according to
the process actions in the order that they appear in the trace. If,
according to the underlying belief-state, the outcome of the next
action scheduled for execution is not predetermined above a threshold
(here 95\%), then the system switches to a DT session.

Because online DT planning is impractical for the size of problem we
are interested in, DT sessions plan in a small abstract problem
defined in terms of the trace from the proceeding sequential
session. This abstract state-space is characterised by a limited
number of propositions, chosen due to their {\em relevance} to the
action whose scheduled execution triggered the switch to the DT
session.  To allow the DT planner to judge assumptions from the trace,
we add {\em disconfirm} and {\em confirm} actions to the problem for
each of them. Those yield a relatively small reward/penalty if the
corresponding judgement is true/false. If a judgement action is
scheduled for execution, then the DT session is terminated, and a new
sequential session begins.

Whatever the session type, our continual planner maintains a factored
representation of successive belief-states.  As an internal
representation of the $(\pp{:init})$ declaration, we keep a
tree-shaped Bayesian network which gets updated whenever an action is
performed, or an observation received. That belief-state
representation is used: (1) as the source of candidate
determinisations for sequential planning, (2) in determining when to
switch to a DT session, and (3) as a mechanism to guide construction
of an abstract process for DT sessions.

\subsection{Sequential Sessions}

As we only consider deterministic-action POMDPs, all state
uncertainty is expressed in the $(\pp{:init})$ declaration. This
declaration is used by our approach to define the starting state for
sequential sessions, and the set of state-assumptions available to
sequential planning.  Writing \#\ if the value of a proposition is
unspecified, taking the $(\pp{:init})$ example from the previous
section, we have the following assumptions:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
\hline
.7 & kitchen & \# &  kitchen\\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
1.0 & kitchen & \# & \# \\
\hline
\end{tabular}
\normalsize

\noindent Each assumption corresponds to one distinct {\em
relaxed} visitation of the root term. Here, a conjunctive term is
visited iff its atomic subterms are visited, and zero or one of its
immediate probabilistic subterms are visited. For a sequential
session, the starting state corresponds to an assumption (usually
unique) with probability $1$.\footnote{Without a loss of generality,
we assume conjunctive terms list atoms that are true with
probability 1 according to a visitation, and that actions do not
have negative preconditions.}  In our example, that starting state is:

\vspace{-1ex}

\small
\[
\begin{array}{l}
\state_0 \equiv \{(=(\pp{is-in}~\pp{R2D2})~\pp{kitchen}),\\
\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
\end{array}
\]
\normalsize

\vspace{-1ex}

To represent state assumptions we augment the problem posed during
a sequential session with an \emph{assumptive action} $\assumptiveS{i}$ for
each element $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. Here, $\assumptiveS{i}$ can be executed if no
$\assumptiveS{j}$, $j \neq i$, has been executed from the same
probabilistic term, and, either
$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
We also add constraints that forbid scheduling of
assumptions about facts after actions with preconditions or effects
that mention those facts. For example, the robot cannot assume it is
plugged into a power source immediately after it unplugs itself.
%%
Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
to a successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$. State
$\state^\bot$ is an added sink.

We now describe the optimisation criteria used during sequential
sessions. Let $\prob_i$ be the probability that the $i^{th}$ sequenced
action, $\action_i$, from a trace of state-action pairs
$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
does not transition to $\state^\bot$, and
$\reward(\state_i, \action_i)$ be the instantaneous reward received
for executing action $\action_i$ in state $\state_i$. The optimal
trace has value:

\small
\[
V^* = \max_N \max_{\state_0, \action_0,.., \state_N} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
\reward(\state_i, \action_i),
\]
\normalsize

\subsection{DT Sessions}

When an action is scheduled whose outcome is uncertain according to
the underlying belief-state, the planner switches to a DT
session. That plans for {\em small} abstract processes defined
according to the action that triggered the DT session, the assumptive
actions in the proceeding trace, and the current
belief-state. Targeted sensing is encouraged by augmenting the reward
model to reflect a heuristic value of knowing the truth about
assumptions. In detail, all rewards from the underlying problem are
retained. Additionally, for each relevant assumptive action
$\assumptiveS{i}$ in the current trace, we have a {\em disconfirm
action} $\assumptiveDT{i}$ so that for all states $\state$:

\vspace{-1ex}
\small
\[
\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
\hat\$(T_i) & \pp{otherwise} \\
\end{array}
\]
\normalsize

\vspace{-1ex}

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a
small positive (negative) numeric quantity which captures
the utility the agent receives for correctly (incorrectly) rejecting
an assumption.
%%
In terms of action physics, a disconfirm action can only be executed
once, and otherwise is modelled as a self-transformation.
%%
We only consider {\em relevant} assumptions when constructing the
abstract model.  If \switchAction\ is the action that switched the
system to a DT session, then an assumption $\assumptiveS{i}$ is {\em
relevant} if it is necessary for the outcome of \switchAction\ to be
determined.  Continuing our simplified example, if the trace is:

\small
\[
\begin{array}{l}
\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{kitchen}));\\
\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{office}));\\
(\pp{look}~\pp{box}~\pp{kitchen});
(\pp{look}~\pp{cup}~\pp{office});\\
(\pp{report}~\pp{box}~\pp{kitchen}); 
(\pp{report}~\pp{cup}~\pp{office})
\end{array}
\]
\normalsize

\noindent Taking the switching action \switchAction\ to be
$(\pp{look}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
relevant for this session, and therefore we exclude the corresponding
disconfirm action from the abstract decision process.

Given \switchAction, we also include another once-only self-transition
action $\actions.\poss(\switchAction)$, a \emph{confirmation action}
with the reward property:

\[
\reward(\state, \actions.\poss(\switchAction)) = \bigg\{ \begin{array}{ll}
\$(\poss(\switchAction)) & \pp{if}~\;\; \poss(\switchAction) \subseteq \state \\
\hat\$(\poss(\switchAction)) & \pp{otherwise} \\
\end{array}
\]

Execution of either a disconfirmation or the confirmation action
returns control to a sequential session, which then continues from the
underlying belief-state.

Turning to the detail of (dis-)confirmation rewards, in our integrated
system these are sourced from a motivational subsystem. In this paper,
for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a small positive
constant, and have $\hat\$(x)= - \$(x)(1 - \prob) /
\prob$ where $\prob$ is the probability that $x$ is true. For
$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
\$(x)\prob/(1-\prob)$.

%%%%

In order to guarantee fast DT sessions, those plan in an abstract
process determined by the current trace and underlying belief-state.
%%
The abstract process posed to the DT planner is constructed by first
constraining as statically false all propositions except those which
are true with probability 1, or which are the subject of {\em
relevant} assumptions. For example, taking the above trace and
switching action {\texttt (look~box~office)}, the underlying
belief in Fig.~\ref{fig:beliefs}B would determine a fully
constrained belief given by Fig.~\ref{fig:beliefs}C.
%%
Next, static constraints are removed, one proposition at a time, until
the number of states that can be true with non-zero probability in the
initial belief of the abstract process reaches a given threshold.
%%
In detail, for each statically-false proposition we compute the {\em
entropy} of the state assumptions of the current trace {\em
conditional} on that proposition. 
%%
Let $X$ be a set of propositions and $2^X$ the powerset of $X$, then taking

\small
\[
\chi = \{\bigwedge_{x \in X'
  \cap X}x \; \land \bigwedge_{x \in X \setminus X'}\neg x \;|\; X' \in 2^X\},
\]
\normalsize

\noindent we have that $\chi$ is a set of conjunctions each of which
corresponds to one truth assignment to elements in $X$. Where
$p(\phi)$ gives the probability that a conjunction $\phi$ holds in the
belief-state of the DTPDDL process, the entropy of $X$
\emph{conditional} on a proposition $y$, written $\entropy(X|y)$, is
given by Eq.~\ref{eq:condent}.

\vspace{-1ex}
\small
\begin{equation}\label{eq:condent}
  \entropy(X|y) = \sum_{x \in \chi, y' \in \{y, \neg y\}} p(x \land y') \log_2
  \frac{p(y')}{p(x \land y')}
\end{equation}
\normalsize

A low $\entropy(X|y)$ value suggests that knowing the truth value of
$y$ is useful for determining whether or not some assumptions $X$
hold. When removing a static constraint on propositions during the
abstract process construction, $y_i$ is considered before $y_j$ if
$\entropy(X|y_i)~<~\entropy(X|y_j)$. For example, if the serial plan
assumes the robot is in a kitchen, then propositions about the
contents of kitchens, e.g. {\texttt (=(is-in~cup)kitchen)}, are
added to characterise the abstract process' states. Taking
assumption $X$ to be {\texttt (=(is-in~box)kitchen)}, in relaxing
static constraints the following entropies are calculated:

\small
\begin{tabtt}
.47 \= = \entropy(\=X|(=(is-in~milk)office))\+ \\
 = \entropy(X|(=(is-in~milk)kitchen)) \-\\
.97 = \entropy(X|(=(is-in~cup)office))\+\\
 = \entropy(X|(=(is-in~cup)kitchen))
\end{tabtt}
\normalsize

\noindent Therefore, the first static constraint to be relaxed is for
\texttt{(=(is-in~milk)office)}, or
equivalently \texttt{(=(is-in~milk)kitchen)}, giving a refined
abstract belief state depicted in
Fig.~\ref{fig:beliefs}A. Summarising, if for Fig.\ref{fig:beliefs}B
the DT session is restricted to belief-states with fewer than $8$
elements, then the starting belief-state of the DT session does not
mention a ``cup''.

\begin{figure}
\tiny
\begin{tabularx}{\columnwidth}{XX}
\hline
\vspace{1ex}
\scriptsize (A) Partially constrained abstract \mbox{\hspace{4ex}belief-state} & 
\vspace{1ex}\scriptsize \hspace{-4ex}(B) Underlying DTPDDL belief \\
\vspace{-2ex} 
\begin{tabtt}
(\=:init (=(is-in~R2D2)kitchen) \+ \\
       (\=.6(and\=(=(is-in~box)kitchen) \+\+ \\
        (\=.9(=(is-in~milk)~kitchen))\+\\
         .1(=(is-in~milk)office))\-\-  \\
       .4(and(=(is-in~box)office) \+\\
        (.1(=(is-in~milk)kitchen))\+\\
        .9(=(is-in~milk)office)))
\end{tabtt} &
\vspace{-2ex} 
\begin{tabtt}
\hspace{-6ex}(\=:init (=(is-in~R2D2)office) \+ \\
       (\=.6(and\=(=(is-in~box)kitchen) \+\+ \\
        (\=.9(=(is-in~milk)kitchen))\+\\
         .1(=(is-in~milk)office))\-\-  \\
       .4(and(=(is-in~box)office) \+\\
        (.1(=(is-in~milk)kitchen))\+\\
        .9(=(is-in~milk)office))) \-\-\-\\
       (\=.6(=(is-in~cup)office) \+ \\
      .4(=(is-in~cup)kitchen))) \\
\end{tabtt} \\
\vspace{-6ex}\scriptsize(C) Fully constrained abstract \mbox{\hspace{5ex}belief-state} & \\
\vspace{-5ex} \begin{tabtt}
(\=:init (=(is-in~R2D2)kitchen) \+ \\
       (\=.6(=(is-in~box)kitchen))) \\
\end{tabtt} & \\
\hline
\end{tabularx}
\caption{Simplified examples of abstract belief-states from
       DT sessions. \label{fig:beliefs}}
\normalsize
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "aaai11"
%%% End: