### 2012-01-20 code/scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex
added comments

\section{Evaluation}\label{sec:evaluation}

To evaluate our progress toward building a cognitive system capable of reasoning about these CDSRs, we conducted the an experiment focusing on the following questions:
\begin{itemize}
\item{How well do anchor points capture context-dependent spatial regions?}
\item{When provided with a base representation containing a labelled CDSR, how well does our approach identify the CDSR in a given target?}
\item{Is the integration of geometric and semantic knowledge necessary?}
\item{Does similarity correlate with performance?}
\end{itemize}

I'm not sure how many of the above questions we want to talk about. The last one is fairly easy and won't take up much space in the paper. The integration of knowledge question is could take some more work as the grouping is determined by semantic type, so in testing just positional knowledge,  I would have to create an alternate set of anchor points.

\subsection{Materials}
We evaluated our approach on six classrooms (two simulated and four real) and two simulated studio apartments. To determine how people define CDRSs, we asked \textbf{SOMENUMBER} of users to draw polygons for particular region types for each room. Users were also shown a polygon drawn by another user and asked to determine if it was an acceptable depiction of the region. All users provided acceptable regions, and therefore, we define the \textit{target region} as the union of the user-defined polygons.

\from{Nick}{What is the base CDSR in a problem instance?}

We consider a \textit{problem instance} to be a room and a sought CDSR type (e.g., ``front''). For each room for which we have a human-labelled target CDSR of the sought type, we generate an \textit{inferred region} using our analogical approach. To assess the quality of the transfer, we use the following measures:

\begin{equation}
	p=\frac{area(inferred region \cap target region)}{area(inferred region)}
\end{equation}
\begin{equation}
	r=\frac{area(inferred region \cap target region)}{area(target region)}
\end{equation}

As a baseline, we will compare our results against the manually encoded CDSRs for the problem instance and the entire room. The precision and recall for considering the entire room provides a baseline from which to assess each of the questions. The performance of the manually encoded CDSRs evaluates how well the anchor points capture the user defined regions and a ceiling for the inferred region performance.

\from{Nick}{What are manually encoded CDSRs? Are these the ones you do and use for the base case?}

\subsection{Results}
\begin{table}
\caption{Performance by Room}
\begin{tabular}{|c|c|c|c|}
\hline
Room & Transfer & Manually Encoded & Entire Room \\
\hline
Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Classroom-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Studio-1 & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
\hline
\end{tabular}
\end{table}

\from{Nick}{Would this be just a single region type or averaging across all applicable ones?}

If we are just going to report means, we should standard deviations as well. %yup!

\begin{table}
\caption{Performance by Region Type}
\begin{tabular}{|c|c|c|c|}
\hline
Region & Transfer & Manually Encoded & Entire Room \\
\hline
Front & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Back & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Front Rows & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Back Rows & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Kitchen & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
Office & p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
\hline
\end{tabular}
\end{table}

Or we could report all of the results in a single entry.

\begin{table}
\caption{Performance against Target Region}
\begin{tabular}{|c|c|c|}
\hline
Transfer & Manually Encoded & Entire Room \\
\hline
p=.6,r=1 & p=.7,r=.9 & p=.7,r=.9 \\
\hline
\end{tabular}
\end{table}

% this would be lovely!

Overall the transferred region was significantly better than the entire room (p<Blah), and not statistically different than the manually encoded region (p=Blah). There was a positive correlation (r=Blah) between the structural evaluation score of the analogy and f-measure of the inference.

\subsection{Discussion}

Our results