### 2012-04-03 code/scenarios/cdsr/trunk/papers/AAAI2012/aaai12-cdsr-rebuttal.txt
added rebuttal text

We thank the reviewers for their great feedback. We would like to
use this space to clarify some aspects of our work. The primary criticism
from Reviewer 2 seems to be that we are not explicitly representing
function/orientation information and just using "spatial templates".

First, to explicitly represent function and recognise it from a mobile robot,
would require a number of advances in computer vision that are outside of our
expertise. It is easier to include orientation information of recognised
objects, and we plan to do so in the future. However object orientation, like object category, is just
a proxy for function (i.e. it represents function implicitly) so it would not
significantly alter our approach, just add more discriminatory power. Rather
than use an explicit representation of function, our approach assumes that
"semantically and geometrically similar areas (e.g. two different classrooms)
will feature similar CDSRs, and that these similarities can be recognised
through analogy." Therefore, we avoid explicitly representing function, but
assume that the functional properties of a region can be extracted through
analogy to a previously labelled (functional) region. If, at some point,
function representations become easily obtainable on a robot, we would be
happy to include them in our system as it should provide a better similarity
measure and assist with analogical transfer.

Second, our use of spatial templates extends only to determining the weights
on different positional relationships between two objects. CDSRs are
represented using anchor points, which are considerably different from
templates. Because anchor points are symbolic, this enables their analogical
transfer which is one of the main contributions of this work. This is a
cognitive systems approach because it allows for new CDSRs to be learned from
a single example. It is not clear how one could use spatial templates to
perform the same task, as the variation across rooms and regions would defeat a template matching approach.

The reviewer raises an important point that there may be gray area in how
these regions are used by people. The results presented in paper are averaged
across each subject. The results by subject were rather similar, although the
regions constructed were a bit different. To address this, we intend to
perform a thorough human study and also embed our system in a navigation
communication task. But the work in the paper shows that our approach can
represent CDSRs and transfer them analogically to semantically and
geometrically similar situations as determined by sensor data.